---
title: "Covid91 vaccine study Final2024"
author: "Alfredo Acevedo"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, covid19 = "covid", OK = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
Male = filter(FinalData, sex =="M")
Female = filter(FinalData, sex =="F")
LGBTQ = filter(FinalData, LGBTQ =="gay")
Druggies = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
  
```{r}
dd2 <- Male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicN <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicN + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position ="dodge")
```
This plot represents the population of males people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 


The plot above shows that a shocking 58% of the male that took the vaccine did in fact get covid-19, while only approximately 42% of the male taking the placebo contracted covid-19.  A substantially higher percentage of males who took tha vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The males that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of males that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the male that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of male who got covid-19, now represents the percentage of male that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of male that got covid-19 after taking the placebo and taking the vaccine. 
Among the males, 58% got covid-19 after taking the vaccine, while the approximately 42% of males that were given the placebo contracted covid-19. The vaccine is clearly not effective in preventing covid-19 in males and it is not recommended. Those who took the placebo were much less likely to contact covid-19
The odds of getting covid-19 while taking the vaccine is 3:2. The males that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
Malefish <- fisher.test(table1)
Malefish
```

The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 6e-04. This is an awfully small number. seeing this number means that these results are not likely to occur by chance.



The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the favorable option.

The odd ratio above illustrates a ratio of 0.68. the males who received the vaccine were 2.4 times more likely to get covid-19 than those who received the placebo. 




```{r}
effect <- Malefish$p.value/(Malefish$p.value +1)
effectivenessMale = Malefish$estimate/(Malefish$estimate +1)*100
print(effectivenessMale)
```

```{r}
low=(Malefish$conf.int[1]/(Malefish$conf.int[1] +1))*100
hi=(Malefish$conf.int[2]/(Malefish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the Covid-19 vaccine for males. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting males from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the males can be seen above as 69%.



## Females

### Barplot

```{r}
dd2 <- Female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of females people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 

In this plot, contrary to the previous, it appears that more female who were only given the placebo contracted covid-19. Only approximately 40% of the individuals who were given the vaccine were reported to have contracted covid-19. That means around 60% of the female that got covid-19 were the individuals that only received the placebo.The odds of getting covid-19 while taking the vaccine is 1:5. The female that took the vaccine were 1.5 less likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of female who got covid-19, now represents the percentage of female that got covid-19 after taking the the placebo or taking the Vaccine. 



The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of females that got covid-19 after taking the placebo and taking the vaccine. 
The females that took the vaccine and contracted covid-19 was 40%, just as approximated in the plots prior. The female that took the placebo and got covid-19 was 60%. The vaccine is in fact effective in protecting female against the contraction of covid-19. 
The odds of getting covid-19 while taking the vaccine is 1:5 for females. The females that took the vaccine were 1.5 times as likely to get covid-19 than those taking the placebo.

### Inferential 




```{r}
chisq.test(table1)
chisqtestGC(table1)
Femalefish <- fisher.test(table1)
Femalefish
```



The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 3e-04. This is an extremely tiny number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.


The Fishers's Exact Test illustrates that the odds ratio is 1.5
Female who got the drug were about 0.20 times more likely to get covid-19 than women that only took the placebo.


```{r}
effect <- Femalefish$p.value/(Femalefish$p.value +1)
effectivenessFemale = Femalefish$estimate/(Femalefish$estimate +1)*100
print(effectivenessFemale)
```

```{r}
low=(Femalefish$conf.int[1]/(Femalefish$conf.int[1] +1))*100
hi=(Femalefish$conf.int[2]/(Femalefish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`




Above is the effectiveness for the covid-19 for females. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we cn conclude that the vaccine is highly effective in protecting females from getting covid-19. 

The hallmark of an effective vaccine is 60% or better. This vaccine works great in protecting females from constructing covid-19.

The odds ratio is approximately how effective the vaccine is. The odds ratio for the female can be seen above as 67. The effectiveness confidence interval for the female is 54.54 to 65.07





## LGBTQ

### Barplot

```{r}
dd2 <- LGBTQ%>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



This plot represents the population of LGBTQ people and the relationship to the contraction of covid-19 by taking either the vaccine or the placebo.

In this barplot, it appears that a sniffling 85% of the individuals that took the vaccine got covid-19 as well as the 15% who received the placebo. The odds of getting covid-19 while taking the vaccine is 2.5:1. The individuals who identified as LGBTQ and took the vaccine were three times as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of LGBTQ individuals that got covid-19 after taking the placebo and taking the vaccine. 
The LGBTQ participants that took the vaccine and contracted covid-19 was at an approximately shocking 85%! Those participants that only received the placebo only contacted covid-19 at 15%. The vaccine id evidently completely ineffective in protecting the LGBTQ community against covid-19 and taking it is not advised. 
The odds of getting covid-19 while taking the vaccine is 0.18. The individuals who identified as LGBTQ and took the vaccine were 0.18 less likely to get covid-19 than those taking the placebo.

### Inferential 




```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```


The p-value is 7e-08. This in a small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 0.18
LGBTQ individuals who got the vaccine were about 0.18 less more likely to get Covid-19 than LGBTQ that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is completely ineffective in preventing the chances of getting covid-19 in individuals who identify as LGBTQ.





```{r}
effect <- LGBTQfish$p.value/(LGBTQfish$p.value +1)
effectivenessLGBTQ = LGBTQfish$estimate/(LGBTQfish$estimate +1)*100
print(effectivenessLGBTQ)
```

```{r}
low=(LGBTQfish$conf.int[1]/(LGBTQfish$conf.int[1] +1))*100
hi=(LGBTQfish$conf.int[2]/(LGBTQfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`


Above is the effectiveness for the covid-19 for those who identified as LGBTQ.  If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting people who identify as LGBTQ from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the LGBTQ folk be seen above as 29. This is completely ineffective. The people who take the vaccine are not better off then those who don't receive it at all. The effectiveness confidence interval for the people who identified as LGBTQ is 7.6 to 26.57.




## Druggies

### Barplot


```{r}
dd2 <-  Druggies %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of people who answered Druggies to drug use and the relationship to the contraction of covid-19 while taking either the vaccine or the placebo.

This barplot shows that around 12% of the people who answered yes to drug use and recieved the vaccine contracted covid-19. This means that around 88% of the druggies that only took the placebo ended up getting covid-19.The odds of getting covid-19 while taking the vaccine is 1:2. The Druggies that took the vaccine were half as likely to get covid-19 than those taking the placebo.




There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 




The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)
colPerc(table1)
```
 

These numerical results show specific percentages of Druggies that got covid-19 after taking the placebo and taking the vaccine.
The percentage of Druggies that took the vaccine and contacted covid-19 is 88% Those that were only given the placebo and got covid-19 was 12%. It is clear that the vaccine to prevent covid-19 is ineffective in protecting Druggies against covid-19 and  so getting this vaccine is not recommended. The odds of getting covid-19 while taking the vaccine is 0.13. The Druggies that took the vaccine were half as likely to get covid-19 than those taking the placebo.

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
Druggiesfish <- fisher.test(table1)
Druggiesfish
```


The p-value is 3e-09. This in a small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 0.13
Druggies who got the vaccine were about 0.13 times as likely to get Covid-19 than Druggies that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is  effective in preventing the chances of getting covid-19 in drugusers.




```{r}
effect <- Druggiesfish$p.value/(Druggiesfish$p.value +1)
effectivenessDruggies = Druggiesfish$estimate/(Druggiesfish$estimate +1)*100
print(effectivenessDruggies)
```

```{r}
low=(Druggiesfish$conf.int[1]/(Druggiesfish$conf.int[1] +1))*100
hi=(Druggiesfish$conf.int[2]/(Druggiesfish$conf.int[2] +1))*100
options(digits=2)
```







## Effectiveness Confidence Interval `r low`  to `r hi`



Above, is the effectiveness for the covid-19 for male. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting Druggies from getting covid-19. 

The hallmark of an effective vaccine is 60% or better. This vaccine works great in protecting Druggies from contracting covid-19.

The odds ratio is approximately how effective the vaccine is. The odds ratio for the Druggies can be seen above as 11. The effectiveness confidence interval for the Druggies is 4.6 to 21.82.


# Overall Results and Conclusions


The covid-19 vaccine was effective in protecting male against the contraction of the disease. The covid-19 effectiveness was `r effectivenessMale`
for Male. It was the most effective in protecting the Males.

The covid-19 vaccine was highly successful in protecting Female against covid-19.The covid-19 effectiveness was `r effectivenessFemale`.

The covid-19 vaccine was not effective in treating people who identified as LGBTQ. The individuals who identified as gay and took the vaccine were two and a half times as likely to get covid-19 than those taking the placebo.The covid-19 effectiveness was `r effectivenessLGBTQ` for LGBTQ individuals.


The drugusers that took the vaccine were 0.13 more likely to get covid-19 than those taking the placebo.The covid-19 effectiveness was `r effectivenessDruggies` for druggies.

If you are a male or female it is recommended to receive the vaccine. However, if you are identifying as LGBTQ or a drug user, the vaccine was reported to have been ineffective.

The null that there is no difference in taking the vaccine can be accepted to the categories including male and female individuals. the alternative, that the vaccine works and is effective, can be accepted for the LGBTQ, and the people who used drugs.


